### 2021-02-27 text/draft.tex
analysis

\begin{figure}
\centering
\includegraphics{images/chi_square_selection_fields.pdf}
\caption{The sample data set (\(n\) = 4'111) is not representative of
the raw Edoc data per department. Data field \texttt{abstract}:
\(\chi^2 (df=9) =\) 3'160.556, \(p < 0.001\); data field
\texttt{id\_number}: \(\chi^2 (df=9) =\) 4'209.0285, \(p < 0.001\); data
field \texttt{keywords}: \(\chi^2 (df=9) =\) 2'314.533, \(p < 0.001\).
The data foundation is available at
\texttt{/edoc/analysis/chi\_square...} and the \(\chi^2\)-statistic can
be calculated by calling \texttt{\_Analysis.get\_chi\_square\_fit} on
the data foundation files.}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{images/raw_sample_analysis.pdf}
\caption{The sample data set (\(n\) = 4'111) is significantly skewed
towards English (data field \texttt{language}: \(\chi^2 (df=3) =\)
1'441.414, \(p < 0.001\)) journal publications (data field
\texttt{type}: \(\chi^2 (df=10) =\) 52'519.743, \(p < 0.001\)) in the
sciences, medicine and economics (data field \texttt{department}:
\(\chi^2 (df=9) =\) 14'447.276, \(p < 0.001\)) from the 21st century
(data field \texttt{date}: \(\chi^2 (df=7) =\) 35'878.493,
\(p < 0.001\)) as compared to the raw Edoc data. The data foundation is
available at \texttt{/edoc/analysis/chi\_square...} and the
\(\chi^2\)-statistic can be calculated by calling
\texttt{\_Analysis.get\_chi\_square\_fit} on the data foundation files.}
\end{figure}

@@

To analyse the spread of keywords in the sample data set, the keywords
per item are counted. To do so, we call \texttt{\_Keywords.count\_all}
on \texttt{/indexed/indexed\_master\_mesh\_enriched.json} and save the
output as \texttt{/analysis/keywords\_counted.json}:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{sample\_data\_set }\OperatorTok{=}\NormalTok{ \_Utility.load\_json(DIR }\OperatorTok{+} \StringTok{"/indexed/indexed\_master\_mesh\_enriched.json"}\NormalTok{)}
\NormalTok{keywords\_counted }\OperatorTok{=}\NormalTok{ \_Keywords.count\_all(sample\_data\_set)}
\NormalTok{\_Utility.save\_json(keywords\_counted, DIR }\OperatorTok{+} \StringTok{"/analysis/keywords\_counted.json"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

Finally, consider the distribution of the cleaned and reconciled
keywords per item in the Edoc sample data set. The corresponding data is
generated with \texttt{\_Keywords.count\_all} as described in
\protect\hyperlink{analysis}{subsection Analysis} above and available as
\texttt{/analysis/keywords\_counted.json}. Figure 5 shows that the
median number of keywords from MeSH or YSO might be too low to be
adequate. This problem can be amended by imposing further constraints on
the sample data set and such a solution is discussed in section {[}!!
section{]}.

@@

Let us briefly look at the definitions and motivations of the chosen
metrics. Recall that a suggestion of a subject term is correct if and
only if the subject term is in the native gold standard. The possible
outcomes are summarized in Table 1.

@@

\(\text{Precision} = \displaystyle \frac{\text{True positive}}{\text{True positive} + \text{False positive}}\)

@@

\(\text{Precision} = \displaystyle \frac{\text{True positive}}{\text{True positive} + \text{False negative}}\)

@@

\(\text{F1} = \displaystyle 2 * \frac{\text{Precision} * \text{Recall}}{\text{Precision} + \text{Recall}}\)

The metric of F1-score is implemented by \texttt{\_Analysis\_get\_f1}

@@

!! how to deal with items that have more than 10 keywords in standard??
just look at data in IQR?

\hypertarget{yso}{%
\paragraph{YSO}\label{yso}}

\hypertarget{wikidata}{%
\paragraph{Wikidata}\label{wikidata}}

@@

>>>>

\begin{figure}
\centering
\includegraphics{images/chi_square_selection_fields.pdf}
\caption{The sample data set (\(n\) = 4'111) is not representative of
the raw Edoc data per department. Data field \texttt{abstract}:
\(\chi^2 (df=9) =\) 3'160.556, \(p < 0.001\); data field
\texttt{id\_number}: \(\chi^2 (df=9) =\) 4'209.0285, \(p < 0.001\); data
field \texttt{keywords}: \(\chi^2 (df=9) =\) 2'314.533, \(p < 0.001\).
The data foundation is available at
\texttt{/edoc/analysis/chi\_square...} and the \(\chi^2\)-statistic can
be calculated by calling \texttt{\_Analysis.print\_chi\_square\_fit} on
the data foundation files.}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{images/raw_sample_analysis.pdf}
\caption{The sample data set (\(n\) = 4'111) is significantly skewed
towards English (data field \texttt{language}: \(\chi^2 (df=3) =\)
1'441.414, \(p < 0.001\)) journal publications (data field
\texttt{type}: \(\chi^2 (df=10) =\) 52'519.743, \(p < 0.001\)) in the
sciences, medicine and economics (data field \texttt{department}:
\(\chi^2 (df=9) =\) 14'447.276, \(p < 0.001\)) from the 21st century
(data field \texttt{date}: \(\chi^2 (df=7) =\) 35'878.493,
\(p < 0.001\)) as compared to the raw Edoc data. The data foundation is
available at \texttt{/edoc/analysis/chi\_square...} and the
\(\chi^2\)-statistic can be calculated by calling
\texttt{\_Analysis.print\_chi\_square\_fit} on the data foundation
files.}
\end{figure}

@@

To analyse the spread of keywords in the sample data set, the keywords
per item are counted. To do so, we call \texttt{\_Keywords.make\_count}
on \texttt{/indexed/indexed\_master\_mesh\_enriched.json} and save the
output as \texttt{/analysis/keywords\_counted.json}:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{sample\_data\_set }\OperatorTok{=}\NormalTok{ \_Utility.load\_json(DIR }\OperatorTok{+} \StringTok{"/indexed/indexed\_master\_mesh\_enriched.json"}\NormalTok{)}
\NormalTok{keywords\_counted }\OperatorTok{=}\NormalTok{ \_Keywords.make\_count(sample\_data\_set)}
\NormalTok{\_Utility.save\_json(keywords\_counted, DIR }\OperatorTok{+} \StringTok{"/analysis/keywords\_counted.json"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

Finally, consider the distribution of the cleaned and reconciled
keywords per item in the Edoc sample data set. The corresponding data is
generated with \texttt{\_Keywords.make\_count} as described in
\protect\hyperlink{analysis}{subsection Analysis} above and available as
\texttt{/analysis/keywords\_counted.json}. Figure 5 shows that the
median number of keywords from MeSH or YSO might be too low to be
adequate. This problem can be amended by imposing further constraints on
the sample data set and such a solution is discussed in section {[}!!
section{]}.

@@

Let us briefly look at the definitions and motivations of the chosen
metrics. Remember that a suggestion of a subject term is correct if and
only if the subject term is in the native gold standard. The possible
outcomes are summarized in Table 1.

@@

\begin{center} 
$\text{Precision} = \displaystyle \frac{\text{True positive}}{\text{True positive} + \text{False positive}}$
\end{center}

@@

\begin{center} 
$\text{Precision} = \displaystyle \frac{\text{True positive}}{\text{True positive} + \text{False negative}}$
\end{center}

@@

\begin{center} 
$\text{F1} = \displaystyle 2 * \frac{\text{Precision} * \text{Recall}}{\text{Precision} + \text{Recall}}$
\end{center}

The metric of F1-score is implemented by \texttt{\_Analysis\_get\_f1}.

@@

In this section, I will assess the perfomance of Annif versus the native
gold standard.

\hypertarget{creating-the-data-foundation}{%
\paragraph{Creating the data
foundation}\label{creating-the-data-foundation}}

I will now describe how the data foundation for the assessment was
created. There are three parameters that need to be distinguished(see
\protect\hyperlink{annif}{section ``Annif''}):

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  The Annif project (algorithm plus vocabulary) responsible for the
  indexing of the sample data set. As per the Annif REST API, these are
  \texttt{yso-en}, \texttt{yso-maui-en}, \texttt{yso-bonsai-en},
  \texttt{yso-fasttext-en}, and \texttt{wikidata-en}.
\item
  The text base per item, namely title versus title and abstract.
\item
  The maximum number of suggestions per item. Since we required 10
  suggestions per item, we can choose between 1-10 suggestions.
\end{enumerate}

Combining these parameters, we really have 100 Annif configurations
whose perfomance we want to compare and assess: Annif project \(*\) text
base \(*\) maximum number of suggestions \(= 5 * 2 * 10 = 100\).

For each configuration, the F1-score was computed for every item in the
sample data set. To do so, we call
\texttt{\_Analysis.super\_make\_metrics} on
\texttt{/indexed/indexed\_master\_mesh\_enriched}. The 100 output files
are saved as \texttt{/metrics/metrics\_\{marker\}} where \texttt{marker}
specifies the Annif configuration:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{\_Analysis.super\_make\_metrics(DIR }\OperatorTok{+} \StringTok{"/indexed/indexed\_master\_mesh\_enriched.json"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

Note that if an item had an empty native gold standard, no score was
computed; this is the case exactly if none of the cleaned keywords had
been matched to Wikidata or YSO respectively. Configurations with a
Wikidata vocabulary had a scoring coverage of 94.38\% of the items in
the sample data set as compared to only 62.84\% for configurations with
a YSO vocabulary.

Next, in order to be able to compare the F1-scores of the 100
configurations, a statistic (quartiles, IQR, mean) was computed for each
configuration by calling \texttt{\_Analysis.super\_make\_stats}, saving
the output as \texttt{/analysis/metrics\_stats.json}:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{\_Analysis.super\_make\_stats()}
\end{Highlighting}
\end{Shaded}

\hypertarget{general-results}{%
\paragraph{General results}\label{general-results}}

we can then discuss whether the best performance is actually useful;
perhaps do this by comparing absolute scores with other assessments. if
results are much lower it is to be suspected that the native gold
standard is not very good (i.e., that the measure we use to judge annif
performance is inadequte =\textgreater{} motivation for foreign
standard)

construct all metrics: \_Analysis.super\_make\_metrics(file\_path=DIR +
``/indexed/indexed\_master\_mesh\_enriched.json'')

construcht all stats for metrics: \_Analysis.super\_make\_stats()

\hypertarget{yso-configurations}{%
\paragraph{YSO configurations}\label{yso-configurations}}

\hypertarget{wikidata-configurations}{%
\paragraph{Wikidata configurations}\label{wikidata-configurations}}

@@

!! but to do this we would first have to map the annif suggested terms
to mesh\ldots! yes.
